---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(robfpca)
```

# Robust PCA for partially observed functional data

This is the R package `robfpca` implementing the robust functional principal component analysis by using the robust mean and covariance function based on Orthogonalized Gnanadesikan-Kettenring (OGK) estimation.


## Installation
```r
# install.packages("devtools")
devtools::install_github("statKim/robfpca")
```

## Example

### Generate partially observed functional data with contamination
First, we generate 100 partially observed curves with 20% of outlying curves.
```{r}
library(robfpca)

# Generate partially observed curves with outliers
set.seed(1)
x <- sim_delaigle(n = 100,
                  type = "partial",
                  out.prop = 0.2,
                  out.type = 1,
                  dist = "normal")
X <- list2matrix(x)   # transform list to matrix
gr <- seq(0, 1, length.out = 51)   # observed timepoints
matplot(gr, t(X), 
        type = "l",
        xlab = "t", ylab = "X(t)")
```


### Robust mean and covariance estimation via OGK estimation
We estimate mean and covariance function using `cov_ogk()` function which implements OGK method via M-estimator.
```{r}
# Robust mean and covariance functions
cov.obj <- cov_ogk(X,
                   type = "huber",
                   smooth = TRUE,
                   bw = 0.1)
mu.ogk <- cov.obj$mean
cov.ogk <- cov.obj$cov
noise.ogk <- cov.obj$noise.var
```

### Competing methods
```{r}
### Yao et al. (2005)
library(fdapace)
bw <- 0.1   # fixed bandwidth
optns <- list(methodXi = "CE", dataType = "Sparse", kernel = "epan", 
              verbose = FALSE, error = FALSE,
              userBwMu = bw, userBwCov = bw)
mu.yao.obj <- GetMeanCurve(Ly = x$Ly, Lt = x$Lt, optns = optns)
cov.yao.obj <- GetCovSurface(Ly = x$Ly, Lt = x$Lt, optns = optns)
mu.yao <- mu.yao.obj$mu
cov.yao <- cov.yao.obj$cov
noise.yao <- 0
```


Following figures show the true and estimated covariance surfaces.
```{r, message=FALSE}
# Covariance surfaces
library(GA)
par(mfrow = c(1, 3))
cov.true <- get_delaigle_cov(gr, model = 2)   # True covariance function
persp3D(gr, gr, cov.true,
        xlab = "s", ylab = "t", zlab = "C(s,t)",
        main = "True",
        theta = -70, phi = 30, expand = 1)
persp3D(gr, gr, cov.yao,
        xlab = "s", ylab = "t", zlab = "C(s,t)",
        main = "Yao et al. (2005)",
        theta = -70, phi = 30, expand = 1)
persp3D(gr, gr, cov.ogk,
        xlab = "s", ylab = "t", zlab = "C(s,t)",
        main = "Proposed",
        theta = -70, phi = 30, expand = 1)
```


### Functional principal component analysis
```{r}
# Yao et al.(2005)
pca.yao.obj <- funPCA(Lt = x$Lt, 
                      Ly = x$Ly,
                      mu = mu.yao, 
                      cov = cov.yao, 
                      sig2 = noise.yao,
                      work.grid = gr,
                      K = 4)

# Proposed
pca.ogk.obj <- funPCA(Lt = x$Lt, 
                      Ly = x$Ly,
                      mu = mu.ogk, 
                      cov = cov.ogk, 
                      sig2 = noise.ogk,
                      work.grid = gr,
                      K = 4)
```

```{r}
# Plot 1~3 eigenfunctions
eig.true <- get_delaigle_eigen(gr, model = 2)   # True eigenfunctions
eig.yao <- check_eigen_sign(pca.yao.obj$eig.fun, 
                            eig.true)   # match eigen directions
eig.ogk <- check_eigen_sign(pca.ogk.obj$eig.fun, 
                            eig.true)   # match eigen directions
par(mfrow = c(1, 3))
for (i in 1:3) {
    plot(gr, eig.true[, i],
         type = "l",
         lwd = 2,
         ylim = c(-3, 3),
         xlab = "t", 
         ylab = paste("PC", i))
    lines(gr, eig.yao[, i], col = 2, lwd = 2)
    lines(gr, eig.ogk[, i], col = 3, lwd = 2)
    
    if (i == 1) {
        legend("bottomright", 
               c("True","Yao et al.(2005)","Proposed"),
               lty = c(1, 1, 1),
               col = c(1, 2, 3))
    }
}
```


### Completion and reconstruction for missing parts
```{r}
# Curve reconstruction
pred_yao_mat <- predict(pca.yao.obj, K = 4)
pred_ogk_mat <- predict(pca.ogk.obj, K = 4)


### Completion and Reconstruction example
i <- 1
NA_ind <- which(is.na(X[i, ]))   # index of missing periods
par(mfrow = c(1, 2))

# Completion
matplot(gr, 
        cbind(X[i, ], 
              pred_missing_curve(X[i, ], pred_yao_mat[i, ], conti = FALSE),
              pred_missing_curve(X[i, ], pred_ogk_mat[i, ], conti = FALSE)),
        type = "l",
        lwd = 2,
        xlab = "t", ylab = "X(t)")
lines(gr[NA_ind], 
      x$x.full[i, NA_ind],
      col = "darkgray", lty = 2, lwd = 2)
grid()
legend("topleft", 
       c("True","Yao et al.(2005)","Proposed"),
       lty = 1:3,
       col = 1:3)

# Reconstruction
matplot(gr, 
        cbind(X[i, ], 
              pred_yao_mat[i, ],
              pred_ogk_mat[i, ]),
        type = "l",
        lwd = 2,
        xlab = "t", ylab = "X(t)")
lines(gr[NA_ind], 
      x$x.full[i, NA_ind],
      col = "darkgray", lty = 2, lwd = 2)
grid()
```


## Reproduced Simulation Results
If you use `Simulation/simulation.R`, you can obtain reproduced results in our paper.


## Reference
<!-- - H. Kim, Y. Park and Y. Lim (2022+). Robust principal component analysis and its applications for partially observed functional data, *Submitted*. -->


